Turned on!": Momentum in Round-Based First Person Shooter Esports

### 2022-07-23 presentation.qmd
re-render pngs with logos, refactor streaks

```{r}
#| label: descriptive-val-snd-stats
#| eval: true
val_n_series <- val_rounds |> 
  distinct(game_id) |> 
  nrow() |> 
  scales::comma()

val_n_rounds <- val_rounds |> 
  distinct(game_id, round) |> 
  nrow() |> 
  scales::comma()

@@

p_val_o_round_win_prop <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  geom_path(
    data = val_pre_swap_round_segments,
    color = 'white',
    size = 1.5,
    aes(
      x = pre_cumu_w, 
      y = pre_cumu_l
    )
  ) +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_round_win_text_layer,
    .threshold = 0.8,
    .labeller = val_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_val_o_round_win_prop
save_plot_for_slide(p_val_o_round_win_prop)
```

@@

```{r}
#| label: p_val_o_series_win_prop
p_val_o_series_win_prop <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  geom_path(
    data = val_pre_swap_round_segments,
    color = 'white',
    size = 1.5,
    aes(
      x = pre_cumu_w, 
      y = pre_cumu_l
    )
  ) +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_series_win_text_layer,
    .threshold = 0.8,
    .labeller = val_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_val_o_series_win_prop
save_plot_for_slide(p_val_o_series_win_prop)
```

@@

val_max_diff_win_series_prop <- max(abs(val_o_win_prop$diff_win_series_prop))
p_val_o_series_win_prop_over_expected <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.2,
    .labeller = val_heatmap_div_labeller
  ) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    limits = c(
      -val_max_diff_win_series_prop,
      +val_max_diff_win_series_prop
    ),
    midpoint = 0
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_val_o_series_win_prop_over_expected
save_plot_for_slide(p_val_o_series_win_prop_over_expected)
```

@@

@@

```{r}
#| label: p_cod_o_win_r1_vs_post_round_prop
calculate_cod_o_round_win_prop <- function(df) {
  df |> 
    count(year, game, map, is_offense, win_round) |> 
    group_by(map, is_offense) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |>
    filter(win_round, is_offense) |>
    select(
      year,
      game,
      map,
      rounds_won = n,
      total_rounds = total,
      win_round_prop = prop
    ) |>
    arrange(desc(win_round_prop))
}

# https://raw.githubusercontent.com/dgrtwo/splittestr/master/R/vectorized-prop-test.R
vectorized_prop_test_approx <- function(a, b, c, d) {
  n1 <- a + b
  n2 <- c + d
  n <- n1 + n2
  p <- (a + c) / n
  E <- cbind(p * n1, (1 - p) * n1, p * n2, (1 - p) * n2)

@@

summarize_cod_o_round_win_prop <- function(.round) {

  specified_round <- cod_rounds |> 
    filter(round == .round) |> 
    calculate_cod_o_round_win_prop()

  post_rounds <- cod_rounds |> 
    filter(round > .round) |> 
    calculate_cod_o_round_win_prop()

  specified_round |> 
    inner_join(
      post_rounds |> rename_with(~sprintf('post_%s', .x), -c(year, game, map)),
      by = c('year', 'game', 'map')
    ) |> 
    mutate(round = .round, .after = 'map') |> 
    arrange(desc(win_round_prop)) |> 
    mutate(
      prop_test = vectorized_prop_test(rounds_won, total_rounds, post_rounds_won, post_total_rounds),
      p_value = prop_test$p.value
    ) |> 
    arrange(p_value) |> 
    mutate(
      diff_prop = post_win_round_prop - win_round_prop,
      dir = factor(sign(diff_prop)),
      is_significant = p_value < 0.05,
      label = sprintf('%s - %s (%s)', map, game, year)
    )
}

@@

plot_cod_round_vs_post_round_slope_chart <- function(.round) {
  o_win_rx_prop <- summarize_cod_o_round_win_prop(.round)
  x1_label <- sprintf('Round %d<br/>offensive win %%', .round)
  x2_label <- sprintf('Post-round %d<br/>offensive win %%', .round)

  min_o_win_rx_prop <- o_win_rx_prop |> slice_min(diff_prop, n = 1)
  max_o_win_rx_prop <- o_win_rx_prop |> slice_max(diff_prop, n = 1)
  minmax_labels <- c(min_o_win_rx_prop$label, max_o_win_rx_prop$label)

  p <- o_win_rx_prop |> 
    ggplot() +
    geom_vline(
      data = tibble(),
      aes(xintercept = c(1, 2)), 
      color = gray_points
    ) +
    geom_hline(
      data = tibble(),
      aes(yintercept = c(0.25, 0.5, 0.75)), 
      linetype = 2,
      color = gray_points
    ) +
    geom_segment(
      data = o_win_rx_prop |> 
        mutate(
          big_segment = label %in% minmax_labels
        ),
      show.legend = FALSE,
      aes(
        x = 1,
        xend = 2,
        y = win_round_prop,
        yend = post_win_round_prop,
        color = dir,
        size = big_segment,
        group = label
      )
    ) +
    geom_text(
      data = tibble(
        y = c(0.25, 0.75)
      ) |> 
        mutate(
          lab = sprintf('%s%%', round(100 * y))
        ),
      aes(x = 2, y = y, label = lab), 
      size = 12 / .pt,
      hjust = -0.05,
      vjust = -0.25,
      fontface = 'bold',
      color = 'white'
    ) +
    ggtext::geom_richtext(
      fill = blackish_background, 
      label.color = NA,
      data = tibble(
        x = c(1, 2),
        y = rep(0.7, 2),
        lab = c(x1_label, x2_label)
      ),
      aes(x = x, y = y, label = lab), 
      size = 9 / .pt,
      hjust = 0.5,
      vjust = 0.5,
      family = font,
      fontface = 'bold',
      color = 'white'
    ) +
    ggrepel::geom_text_repel(
      data = o_win_rx_prop |> 
        filter(label %in% !!minmax_labels),
      aes(x = 1, y = win_round_prop, label = label),
      family = font,
      fontface = 'bold',
      color = 'white',
      size = 9 / .pt,
      hjust = 'right', 
      nudge_x = -0.1,
      direction = 'y'
    ) +
    ggrepel::geom_text_repel(
      data = o_win_rx_prop |> 
        filter(!(label %in% !!minmax_labels)),
      aes(x = 2, y = post_win_round_prop, label = label),
      family = font,
      color = gray_text,
      size = 9 / .pt,
      hjust = 'left', 
      nudge_x = 0.1,
      direction = 'y'
    ) +
    scale_x_continuous(
      position = 'top',
      breaks = c(1, 2),
      expand = expansion(mult = 0.5)
    ) +
    scale_y_continuous(
      limits = c(0.25, 0.75)
    ) +
    scale_size_manual(
      values = c(
        `FALSE` = 0.5,
        `TRUE` = 2
      )
    ) +
    scale_color_manual(
      name = NULL,
      values = o_win_rx_prop_pal
    ) +
    theme(
      plot.title = element_text(hjust = 0.5),
      panel.grid.major = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank()
    ) +
    labs(
      x = NULL,
      y = NULL
    )
  list(
    plot = p,
    data = o_win_rx_prop,
    largest_neg_diff_prop = min_o_win_rx_prop$diff_prop,
    largest_pos_diff_prop = max_o_win_rx_prop$diff_prop
  )
}

cod_res_r1 <- plot_cod_round_vs_post_round_slope_chart(1)

p_cod_o_win_r1_vs_post_round_prop <- cod_res_r1$plot +
  geom_curve(
    data = tibble(),
    aes(x = 1.2, y = 0.65, xend = 1.1, yend = 0.63),
    curvature = 0.2, 
    color = o_win_rx_prop_pal[["-1"]], 
    arrow = arw
  ) +
  geom_curve(
    data = tibble(),
    aes(x = 1.24, y = 0.27, xend = 1.1, yend = 0.27),
    curvature = 0-.2, 
    color = o_win_rx_prop_pal[["1"]], 
    arrow = arw
  ) +
  geom_text(
    data = tibble(
      x = c(1.2, 1.24),
      y = c(0.65, 0.27),
      hjust = c(0, 0),
      vjust = c(0.5, 0.5),
      lab = c(
        sprintf(
          'Largest negative percent difference in round 1\nvs. all other round win percentages (%+.0f%%).', 
          100 * cod_res_r1$largest_neg_diff_prop
        ),
        sprintf(
          'Largest positive percent difference (%+.0f%%).\nOnly significant difference (p value < 0.01).',
          100 * cod_res_r1$largest_pos_diff_prop
        )
      )
    ),
    color = 'white',
    size = 9 / .pt,
    aes(x = x, y = y, hjust = hjust, vjust = vjust, label = lab)
  )
p_cod_o_win_r1_vs_post_round_prop
save_plot_for_slide(p_cod_o_win_r1_vs_post_round_prop)
```

@@

@@

```{r}
#| label: p_cod_series_outcome_prop_over_expected
# https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422
prob_of_series_lasting_n_games <- function(n, p = 0.5, max_rounds = 11) {
  s <- ceiling(max_rounds / 2)
  (factorial(n - 1) / (factorial(s - 1) * factorial(n - s))) * (p^s * (1 - p)^(n - s) + p^(n - s) * (1 - p)^s)
}

possibilities_of_outcomes <- function(m, n) {
  factorial(m + n) / (factorial(m) * factorial(n))
}

cod_actual_series_streaks <- cod_rounds |> 
  filter(win_series) |> 
  mutate(across(win_round, as.integer)) |> 
  group_by(year, event, series) |> 
  summarize(
    wins = max(cumu_w),
    losses = max(cumu_l),
    ws = paste0(win_round, collapse = '')
  ) |> 
  ungroup() |> 
  mutate(n_rounds = wins + losses) |> 
  unite(
    record, wins, losses, sep = '-'
  ) |> 
  count(n_rounds, record, ws, sort = TRUE) |> 
  mutate(prop = n / sum(n))

cod_possibilities <- tibble(
  n_rounds = 6L:11L
) |> 
  mutate(
    series_prop = map_dbl(n_rounds, prob_of_series_lasting_n_games),
    possibilities = map2_dbl(6, n_rounds - 6, possibilities_of_outcomes)
  ) |> 
  transmute(
    n_rounds,
    series_prop,
    record = sprintf('6-%d', n_rounds - 6),
    prop = series_prop / possibilities
  )

cod_series_streaks <- inner_join(
  cod_actual_series_streaks |> 
    rename_with(~sprintf('%s_actual', .x), c(n, prop)),
  cod_possibilities |> 
    rename_with(~sprintf('%s_expected', .x), prop),
  by = 'n_rounds'
) |> 
  mutate(
    prop_diff = prop_actual - prop_expected
  ) |> 
  arrange(desc(abs(prop_diff)))
# series_streaks |> filter(is.na(n_actual))

## 6-0, 6-1, and some combos of 6-5 occur more often than expected
cod_series_streak_props <- cod_series_streaks |> 
  filter(!is.na(n_actual)) |> 
  mutate(
    prop_diff = prop_actual - prop_expected,
    total_actual = sum(n_actual),
    n_expected = round(prop_expected * total_actual),
    p = vectorized_prop_test(n_actual, total_actual, n_expected, total_actual)
  ) |> 
  select(-total_actual) |> 
  unnest_wider(p) |> 
  arrange(p.value)

## 6-0 and 6-1 occur more often than expected
cod_series_outcome_prop <- full_join(
  cod_actual_series_streaks |> 
    group_by(record) |> 
    summarize(
      across(n, sum)
    ) |> 
    mutate(prop = n / sum(n)) |> 
    rename_with(~sprintf('%s_actual', .x), c(n, prop)),
  cod_possibilities |> 
    select(record, prop_expected = series_prop),
  by = 'record'
) |> 
  mutate(
    prop_diff = prop_actual - prop_expected,
    total_actual = sum(n_actual),
    n_expected = round(prop_expected * total_actual),
    p = vectorized_prop_test(n_actual, total_actual, n_expected, total_actual)
  ) |> 
  select(-total_actual) |> 
  unnest_wider(p) |> 
  arrange(p.value)

cod_long_series_outcome_prop <- cod_series_outcome_prop |>
  select(
    record,
    n = n_actual,
    `3` = prop_actual,
    `2` = prop_expected,
    `1` = prop_diff
  ) |> 
  pivot_longer(
    matches('[123]'),
    names_to = 'outcome_idx',
    values_to = 'value'
  ) |> 
  mutate(across(outcome_idx, as.integer))

>>>>

```{r}
#| label: descriptive-val-snd-stats
#| eval: true
val_n_series <- val_rounds |> 
  distinct(series_id) |> 
  nrow() |> 
  scales::comma()

val_n_rounds <- val_rounds |> 
  distinct(series_id, round) |> 
  nrow() |> 
  scales::comma()

@@

val_swap_path <- function(...) {
  list(
    ...,
    geom_path(
      data = val_pre_swap_round_segments,
      color = 'white',
      size = 1.5,
      aes(
        x = pre_cumu_w, 
        y = pre_cumu_l
      )
    )
  )
}

p_val_o_round_win_prop <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  val_swap_path() +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_round_win_text_layer,
    .threshold = 0.8,
    .labeller = val_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_val_o_round_win_prop
save_plot_for_slide(p_val_o_round_win_prop)
```

@@

```{r}
#| label: p_val_o_series_win_prop
p_val_o_series_win_prop <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  val_swap_path() +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_series_win_text_layer,
    .threshold = 0.8,
    .labeller = val_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_val_o_series_win_prop
save_plot_for_slide(p_val_o_series_win_prop)
```

@@

val_max_diff_win_series_prop <- max(abs(val_o_win_prop$diff_win_series_prop))
p_val_o_series_win_prop_over_expected <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.2,
    .labeller = val_heatmap_div_labeller
  ) +
  val_swap_path() +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    limits = c(
      -val_max_diff_win_series_prop,
      +val_max_diff_win_series_prop
    ),
    midpoint = 0
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_val_o_series_win_prop_over_expected
save_plot_for_slide(p_val_o_series_win_prop_over_expected)
```

@@

```{r}
#| label: p_filt_val_o_series_win_prop_over_expected
filt_val_o_win_prop <- val_o_win_prop |> 
  mutate(
    across(
      diff_win_series_prop,
      ~case_when(
        pre_cumu_w <= 3 & pre_cumu_l <= 3 ~ NA_real_,
        TRUE ~ .x
      )
    )
  )
filt_val_max_diff_win_series_prop <- max(abs(filt_val_o_win_prop$diff_win_series_prop))

p_filt_val_o_series_win_prop_over_expected <- filt_val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  # geom_abline(
  #   color = 'white',
  #   linetype = 2,
  #   aes(
  #     slope = 1,
  #     intercept = 0
  #   )
  # ) +
  heatmap_text_layers(
    .df = filt_val_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.1,
    .labeller = val_heatmap_div_labeller
  ) +
  val_swap_path() +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    limits = c(
      -filt_val_max_diff_win_series_prop,
      +filt_val_max_diff_win_series_prop
    ),
    midpoint = 0
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_filt_val_o_series_win_prop_over_expected
save_plot_for_slide(p_val_o_series_win_prop_over_expected)
```

@@

```{r}
#| label: p_cod_o_win_r1_vs_post_round_prop
# https://raw.githubusercontent.com/dgrtwo/splittestr/master/R/vectorized-prop-test.R
vectorized_prop_test_approx <- function(a, b, c, d) {
  n1 <- a + b
  n2 <- c + d
  n <- n1 + n2
  p <- (a + c) / n
  E <- cbind(p * n1, (1 - p) * n1, p * n2, (1 - p) * n2)

@@

calculate_map_o_round_win_prop <- function(df) {
  df |> 
    count(map, is_offense, win_round) |> 
    group_by(map, is_offense) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |>
    filter(win_round, is_offense) |>
    select(
      map,
      rounds_won = n,
      total_rounds = total,
      win_round_prop = prop
    ) |>
    arrange(desc(win_round_prop))
}

summarize_map_o_round_win_prop <- function(df, .round) {

  specified_round <- df |> 
    filter(round == .round) |> 
    calculate_map_o_round_win_prop()

  post_rounds <- df |> 
    filter(round > .round) |> 
    calculate_map_o_round_win_prop()

  specified_round |> 
    inner_join(
      post_rounds |> rename_with(~sprintf('post_%s', .x), -c(map)),
      by = 'map'
    ) |> 
    mutate(round = .round, .after = 'map') |> 
    arrange(desc(win_round_prop)) |> 
    mutate(
      prop_test = vectorized_prop_test(rounds_won, total_rounds, post_rounds_won, post_total_rounds),
      p_value = prop_test$p.value
    ) |> 
    arrange(p_value) |> 
    mutate(
      diff_prop = post_win_round_prop - win_round_prop,
      dir = factor(sign(diff_prop)),
      is_significant = p_value < 0.05
    )
}

@@

plot_round_vs_post_round_slope_chart <- function(df, .round, game = c('cod', 'val')) {
  game <- match.arg(game)
  o_win_rx_prop <- summarize_map_o_round_win_prop(df, .round)
  x1_label <- sprintf('Round %d<br/>offensive win %%', .round)
  x2_label <- sprintf('Post-round %d<br/>offensive win %%', .round)

  min_o_win_rx_prop <- o_win_rx_prop |> slice_min(diff_prop, n = 1)
  max_o_win_rx_prop <- o_win_rx_prop |> slice_max(diff_prop, n = 1)
  minmax_maps <- c(min_o_win_rx_prop$map, max_o_win_rx_prop$map)

  is_cod <- game == 'cod'
  if (is_cod) {
    y_lo <- 0.25
    y_hi <- 0.75
  } else {
    y_lo <- 0.45
    y_hi <- 0.55
  }

  p <- o_win_rx_prop |> 
    ggplot() +
    geom_vline(
      data = tibble(),
      aes(xintercept = c(1, 2)), 
      color = gray_points
    ) +
    geom_hline(
      data = tibble(),
      aes(yintercept = c(y_lo, 0.5, y_hi)), 
      linetype = 2,
      color = gray_points
    ) +
    geom_segment(
      data = o_win_rx_prop |> 
        mutate(
          big_segment = map %in% minmax_maps
        ),
      show.legend = FALSE,
      aes(
        x = 1,
        xend = 2,
        y = win_round_prop,
        yend = post_win_round_prop,
        color = dir,
        size = big_segment,
        group = map
      )
    ) +
    geom_text(
      data = tibble(
        y = c(y_lo, y_hi)
      ) |> 
        mutate(
          lab = sprintf('%s%%', round(100 * y))
        ),
      aes(x = 2, y = y, label = lab), 
      size = 12 / .pt,
      hjust = -0.05,
      vjust = -0.25,
      fontface = 'bold',
      color = 'white'
    ) +
    ggtext::geom_richtext(
      fill = blackish_background, 
      label.color = NA,
      data = tibble(
        x = c(1, 2),
        y = rep(0.7, 2),
        lab = c(x1_label, x2_label)
      ),
      aes(x = x, y = y, label = lab), 
      size = 9 / .pt,
      hjust = 0.5,
      vjust = 0.5,
      family = font,
      fontface = 'bold',
      color = 'white'
    ) +
    ggrepel::geom_text_repel(
      data = o_win_rx_prop |> 
        filter(map %in% !!minmax_maps),
      aes(x = 1, y = win_round_prop, label = map),
      family = font,
      fontface = 'bold',
      color = 'white',
      size = 9 / .pt,
      hjust = 'right', 
      nudge_x = -0.1,
      direction = 'y'
    ) +
    ggrepel::geom_text_repel(
      data = o_win_rx_prop |> 
        filter(!(map %in% !!minmax_maps)),
      aes(x = 2, y = post_win_round_prop, label = map),
      family = font,
      color = gray_text,
      size = 9 / .pt,
      hjust = 'left', 
      nudge_x = 0.1,
      direction = 'y'
    ) +
    scale_x_continuous(
      position = 'top',
      breaks = c(1, 2),
      expand = expansion(mult = 0.5)
    ) +
    scale_y_continuous(
      limits = c(y_lo, y_hi)
    ) +
    scale_size_manual(
      values = c(
        `FALSE` = 0.5,
        `TRUE` = 2
      )
    ) +
    scale_color_manual(
      name = NULL,
      values = o_win_rx_prop_pal
    ) +
    theme(
      plot.title = element_text(hjust = 0.5),
      panel.grid.major = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank()
    ) +
    labs(
      x = NULL,
      y = NULL
    )
  list(
    plot = p,
    data = o_win_rx_prop,
    largest_neg_diff_prop = min_o_win_rx_prop$diff_prop,
    largest_pos_diff_prop = max_o_win_rx_prop$diff_prop
  )
}

cod_res_r1 <- cod_rounds |> 
  mutate(
    map = sprintf('%s - %s (%s)', map, game, year)
  ) |> 
  plot_round_vs_post_round_slope_chart(1, game = 'cod')

p_cod_o_win_r1_vs_post_round_prop <- cod_res_r1$plot +
  geom_curve(
    data = tibble(),
    aes(x = 1.2, y = 0.65, xend = 1.1, yend = 0.63),
    curvature = 0.2, 
    color = o_win_rx_prop_pal[["-1"]], 
    arrow = arw
  ) +
  geom_curve(
    data = tibble(),
    aes(x = 1.24, y = 0.27, xend = 1.1, yend = 0.27),
    curvature = 0-.2, 
    color = o_win_rx_prop_pal[["1"]], 
    arrow = arw
  ) +
  geom_text(
    data = tibble(
      x = c(1.2, 1.24),
      y = c(0.65, 0.27),
      hjust = c(0, 0),
      vjust = c(0.5, 0.5),
      lab = c(
        sprintf(
          'Largest negative percent difference in round 1\nvs. all other round win percentages (%+.0f%%).', 
          100 * cod_res_r1$largest_neg_diff_prop
        ),
        sprintf(
          'Largest positive percent difference (%+.0f%%).\nOnly significant difference (p value < 0.01).',
          100 * cod_res_r1$largest_pos_diff_prop
        )
      )
    ),
    color = 'white',
    size = 9 / .pt,
    aes(x = x, y = y, hjust = hjust, vjust = vjust, label = lab)
  ) +
  common_cod_labs()
p_cod_o_win_r1_vs_post_round_prop
save_plot_for_slide(p_cod_o_win_r1_vs_post_round_prop)
```

@@

## Are offensive win %'s different in round 1 ([Valorant]{.valorant}?

```{r}
#| label: p_val_o_win_r1_vs_post_round_prop
p_val_o_win_r1_vs_post_round_prop <- val_rounds |> 
  plot_round_vs_post_round_slope_chart(1, game = 'val') |> 
  pluck('plot') +
  common_val_labs()
save_plot_for_slide(p_val_o_win_r1_vs_post_round_prop)
```

![](val_o_win_r1_vs_post_round_prop.png){fig-align="center"}

@@

```{r}
#| label: p_cod_series_outcome_prop_over_expected
# https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422
prob_of_series_lasting_n_games <- function(n, p = 0.5, max_round) {
  s <- ceiling(max_round / 2)
  (factorial(n - 1) / (factorial(s - 1) * factorial(n - s))) * (p^s * (1 - p)^(n - s) + p^(n - s) * (1 - p)^s)
}

expected_series_streaks_of_outcomes <- function(m, n) {
  factorial(m + n) / (factorial(m) * factorial(n))
}

summarize_streaks <- function(rounds, game = c('cod', 'val')) {
  rounds <- val_rounds
  game = 'val'
  .max_round <- ifelse(game == 'cod', 11, 25)

  sudden_death_rounds <- rounds |> 
    filter(max_round > .max_round)

  rounds1 <- rounds
  ## should only happen in val
  if(nrow(sudden_death_rounds) > 0) {

    sudden_death_series_ids <- sudden_death_rounds |> 
      distinct(series_id)

    imputed_suddent_death_rounds <- sudden_death_rounds |> 
      mutate(
        round = .max_round
      ) |> 
      select(series_id, round, is_offense) |> 
      inner_join(
        rounds,
        by = c('series_id', 'round', 'is_offense')
      )

    rounds <- bind_rows(
      rounds |> 
        filter(max_round <= .max_round),
      imputed_suddent_death_rounds
    )
  }

  cutoff <- (.max_round + 1) / 2

  actual_round_streaks <- rounds |> 
    filter(round <= .max_round) |> 
    filter(win_series) |> 
    mutate(across(win_round, as.integer)) |> 
    group_by(series_id) |> 
    summarize(
      wins = max(cumu_w),
      losses = max(cumu_l),
      ws = paste0(win_round, collapse = '')
    ) |> 
    ungroup() |> 
    mutate(n_rounds = wins + losses) |> 
    unite(
      record, wins, losses, sep = '-'
    ) |>
    count(record, n_rounds, ws, sort = TRUE) |> 
    mutate(prop = n / sum(n))

  expected_round_streaks <- tibble(
    n_rounds = cutoff:.max_round
  ) |> 
    mutate(
      series_prop = map_dbl(n_rounds, ~prob_of_series_lasting_n_games(.x, max_round = .max_round)),
      n_expected_series_streaks = map2_dbl(cutoff, n_rounds - cutoff, expected_series_streaks_of_outcomes)
    ) |> 
    transmute(
      n_rounds,
      series_prop,
      prop = series_prop / n_expected_series_streaks
    )

  round_streaks <- full_join(
    actual_round_streaks |> 
      rename_with(~sprintf('%s_actual', .x), c(n, prop)),
    expected_round_streaks |> 
      rename_with(~sprintf('%s_expected', .x), prop),
    by = 'n_rounds'
  )

  round_streak_prop <- round_streaks |> 
    drop_na() |> 
    mutate(
      prop_diff = prop_actual - prop_expected,
      total_actual = sum(n_actual),
      n_expected = round(prop_expected * total_actual),
      p = vectorized_prop_test(n_actual, total_actual, n_expected, total_actual)
    ) |> 
    select(-total_actual) |> 
    unnest_wider(p) |> 
    arrange(p.value)

  series_outcomes <- full_join(
    actual_round_streaks |> 
      group_by(record, n_rounds) |> 
      summarize(
        across(n, sum)
      ) |> 
      ungroup() |> 
      mutate(prop = n / sum(n)) |> 
      rename_with(~sprintf('%s_actual', .x), c(n, prop)),
    expected_round_streaks |> 
      select(n_rounds, prop_expected = series_prop),
    by = 'n_rounds'
  )

  series_outcome_prop <- series_outcomes |> 
    drop_na() |> 
    mutate(
      prop_diff = prop_actual - prop_expected,
      total_actual = sum(n_actual),
      n_expected = round(prop_expected * total_actual),
      p = vectorized_prop_test(n_actual, total_actual, n_expected, total_actual)
    ) |> 
    select(-total_actual) |> 
    unnest_wider(p) |> 
    arrange(p.value)

  list(
    rounds = series_streak_prop,
    series = series_outcome_prop
  )
}

cod_streaks_res <- cod_rounds |> summarize_streaks(game = 'cod')
val_streaks_res <- val_rounds |> summarize_streaks(game = 'val')
cod_long_series_outcome_prop <- cod_streaks_res$series |>
  select(
    record,
    n = n_actual,
    `3` = prop_actual,
    `2` = prop_expected,
    `1` = prop_diff
  ) |> 
  pivot_longer(
    matches('[123]'),
    names_to = 'outcome_idx',
    values_to = 'value'
  ) |> 
  mutate(across(outcome_idx, as.integer))
